---
title: "Day 6 - stringr::str_trim()"
description: |
  The 6th most used R package is... stringr!
date: 01-26-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=F, warning=F)
library(tidyverse)

TopRpack <- read.csv("https://github.com/v-kozhevnikov/GitHub_R_commands/raw/master/data/top_100_packages.csv")
TopRFunc <- read.csv("https://github.com/v-kozhevnikov/GitHub_R_commands/raw/master/data/top_2000_functions.csv")
```

### stringr

the `stringr` package is all about the manipulation of strings in R (pretty self explanatory right?). It's inclusion in the tidyverse collective means, in my mind, it is one of the most useful packages for R out there. I haven't used `stringr` much, but I am familiar with `regex` or regular expressions using python to search for and replace specific patterns of strings. By the looks of it, `stringr` makes the syntax for string detection and manipulation a little easier in R and it works out of the box with `%>%` function chains. The regex functionality for more advanced string manipulations is also available using `stringr`.

For more `stringr` introductions: [stringr vignette](https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html)

```{r echo = F, results = 'show', preview = T}
TopRFunc.stringr <- read.csv("https://github.com/v-kozhevnikov/GitHub_R_commands/raw/master/data/top_2000_functions.csv") %>%
  filter(library == 'stringr')

knitr::kable(TopRFunc.stringr[1:10,2:3],
             table.attr = "style='width:60%;'",
             col.names = Hmisc::capitalize(names(TopRFunc.stringr[,2:3])),
             caption = "Top 10 stringr Functions") %>%
  kableExtra::kable_styling(position = "center",
                            bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### stringr::str_trim()

More often then not, a split string or a hand-me-down data set will include some strings with strange and difficult to visually detect white space such as accidental tab spaces, extra spaces, and return lines. `str_trim()` helps remove all leading and trailing white space from strings. The function itself is easy to remember and always useful. Similarly, `str_squish()` removes excess white space.

#### Example

```{r}
str_trim("  String with trailing and leading white space\t")
str_trim("\n\nString with trailing and leading white space\n\n")

str_squish("  String with trailing,  middle, and leading white space\t")
str_squish("\n\nString with excess,  trailing and leading white   space\n\n")
```


