---
title: "Day 1 - ggplot2::qplot()"
description: |
  The most used R package is... ggplot! Surprise?
date: 01-22-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=F, warning=F)
library(tidyverse)

TopRpack <- read.csv("https://github.com/v-kozhevnikov/GitHub_R_commands/raw/master/data/top_100_packages.csv")
TopRFunc <- read.csv("https://github.com/v-kozhevnikov/GitHub_R_commands/raw/master/data/top_2000_functions.csv")
```

Emily Nebergall started this challenge for our Data Science Discord group. The goal is to learn a new R function every day and share what we have learned! R is a big code universe. I started today by perusing a "big data" analysis of the [top 100 R packages](https://towardsdatascience.com/top-100-most-used-r-functions-on-github-9caf2b81b314) used on all of GitHub. Each day, I will move down the list and pick a function I am unfamiliar with from the most popular [R packages](https://github.com/v-kozhevnikov/GitHub_R_commands/blob/master/data/top_100_packages.csv).
I've used `ggplot2` pretty extensively so finding something unfamiliar to me was challenging. Let's see what the top 10 or 20 `ggplot` functions are:
<br>

```{r results = 'show'}
TopRFunc.ggplot <- read.csv("https://github.com/v-kozhevnikov/GitHub_R_commands/raw/master/data/top_2000_functions.csv") %>%
  filter(library == 'ggplot2')

knitr::kable(TopRFunc.ggplot[1:20,2:3],
             table.attr = "style='width:60%;'",
             col.names = Hmisc::capitalize(names(TopRFunc.ggplot[,2:3])),
             caption = "Top 20 ggplot2 Functions on GitHub") %>%
  kableExtra::kable_styling(position = "center",
                            bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```
<br>

### ggplot2::qplot
No surprise still with the top 10 or even 20 ggplot functions. I had to scroll down to **#45** (not bad!) before finding one I haven't heard of: [**quick plot!**](https://ggplot2.tidyverse.org/reference/qplot.html)

`qplot` is meant to be a shortcut for plotting data. It lacks the fine control ggplot has but it works great for quickly visualizing simple data sets.
<br>

#### Example
```{r}
qplot(body_mass_g, flipper_length_mm, data = palmerpenguins::penguins, color = species)
```
<br>

