---
title: "Day 3 - plyr::ddply()"
description: |
  The 3rd most used R package is... plyr
date: 01-24-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=F, warning=F)
library(tidyverse)

TopRpack <- read.csv("https://github.com/v-kozhevnikov/GitHub_R_commands/raw/master/data/top_100_packages.csv")
TopRFunc <- read.csv("https://github.com/v-kozhevnikov/GitHub_R_commands/raw/master/data/top_2000_functions.csv")
```

### plyr!

`plyr` is not a package I've had to use frequently – likely because I use other tidyverse packages. `dplyr` was meant to be faster and more efficient than `plyr`. Perhaps `plyr` still has some uses. Here are the top 10 `plyr` functions:

```{r}
TopRFunc.plyr <- read.csv("https://github.com/v-kozhevnikov/GitHub_R_commands/raw/master/data/top_2000_functions.csv") %>%
  filter(library == 'plyr')

knitr::kable(TopRFunc.plyr[1:10,2:3],
             table.attr = "style='width:60%;'",
             col.names = Hmisc::capitalize(names(TopRFunc.plyr[,2:3])),
             caption = "Top 10 plyr Functions on GitHub") %>%
  kableExtra::kable_styling(position = "center",
                            bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```
<br>

### plyr::ddply
$#1$ on our function list, `plyr`'s [ddply](https://www.rdocumentation.org/packages/plyr/versions/1.8.6/topics/ddply). `ddply` takes a tedious task and makes it fit in one line:
<br><br>
```Split a Data Frame, Apply a Function, And Return Results In A Data Frame.```
<br>

#### Example with Iris Data

```{r}
# Borrowed from https://stackoverflow.com/questions/54774280/plyrddply-equivalent-in-dplyr
# A function
func = function(x) {
  mod = lm(Sepal.Length ~ Petal.Width, data = x)
  mod_coefs = broom::tidy(mod)
}

# plyr::ddply used with predefined function
plyr.df <- plyr::ddply(iris, "Species", func)
plyr.df
```
<br>

I am more familiar with `dplyr` methods that would look like this:

```{r}
# dplyr function chain style
dplyr.df <- iris %>%
  group_by(Species) %>%
  do(func(.)) # uses grouped iris above as input
dplyr.df
```
